---
title: "Data Wrangling for CITES"
author: "Ali Martin and Sarah Lam"
date: "2/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(here)
library(plotly)
library(highcharter)
```


read in the data 
```{r}
# import data 
wildlife <- read_csv(here("data","cites_trade_data.csv")) %>% 
  clean_names()
```

### Wrangling for Widget 1 

import export data 

```{r}
import_sum <- wildlife %>% 
  group_by(importer) %>% 
  summarize(import_count = n()) %>% 
  mutate(code = importer)

export_sum <- wildlife %>% 
  group_by(exporter) %>% 
  summarize(export_count = n()) %>% 
  mutate(code = exporter)
```

wrangling for map
```{r}
country_codes <- read_csv('https://raw.githubusercontent.com/sdlam/ISO-3166-Countries-with-Regional-Codes/master/slim-2/slim-2.csv') %>% 
  select(country = name, code = 'alpha-2') 

world_sf <- read_sf(here('ne_50m_admin_0_countries', 'ne_50m_admin_0_countries.shp'))

world_subset_sf <- world_sf %>% 
  clean_names() %>% 
  select(country = sovereignt) %>% 
  merge(country_codes, by = 'country')

world_import_sf <- merge(world_subset_sf, import_sum, by = 'code')
import_export_sf <- merge(world_import_sf, export_sum, by = 'code')
  #pivot_longer(cols = c("import_count", "export_count"))

ggplot(data = import_export_sf) +
  geom_sf(aes(fill = import_count), color = 'white', size = 0.1) +
  scale_fill_gradient() +
  theme_void()

ggplot(data = import_export_sf) +
  geom_sf(aes(fill = export_count), color = 'white', size = 0.1) +
  scale_fill_gradient() +
  theme_void()


```

```{r}
## map of exports
tmap_mode(mode = 'view') #toggles to interactive viewing
tm_shape(import_export_sf) + #taking a shape..
  tm_fill(col = 'import_count', 
          palette = "viridis",
          alpha = 0.5,
          popup.vars = c("country" = "country", "count of imports" = "import_count")) + #attributing a fill to that shape
  tm_borders(col = 'black')

## top 5 import countries: Japan (JP), USA (US), China (CN), Canada (CA), France (FR)
```


```{r}
## map of exports
tmap_mode(mode = 'view') #toggles to interactive viewing
tm_shape(import_export_sf) + #taking a shape..
  tm_fill(col = 'export_count', 
          palette = "viridis",
          alpha = 0.5,
          popup.vars = c("country" = "country", "count of exports" = "export_count")) + #attributing a fill to that shape
  tm_borders(col = 'black')

```

```{r}
import_countries <- wildlife %>% 
  filter(importer %in% c('JP','US','CN','CA','FR')) %>% 
  group_by(importer) %>% 
count(taxon)


```


```{r}
#count number of diff species imported by country
species_imports <- wildlife %>% 
  select(taxon, importer, year) %>%
  group_by(importer) %>% 
  count(taxon) %>% 
  mutate(code = importer)
  

#do the same for export
species_exports <- wildlife %>% 
  select(taxon, exporter, year) %>%
  group_by(exporter) %>% 
  count(taxon) %>% 
  mutate(code = exporter)

#convert to sf
species_imports_sf <- merge(world_subset_sf, species_imports, by = "code")
species_exports_sf <- merge(world_subset_sf, species_exports, by = "code")

## filter because the data set is too large
top_species_imports <- species_imports_sf %>% 
  filter(taxon %in% c('Oryx dammah','Loxodonta africana','Python reticulatus','Crocodylus niloticus'))

## do the same for exports
top_species_exports <- species_exports_sf %>% 
  filter(taxon %in% c('Oryx dammah','Loxodonta africana','Python reticulatus','Crocodylus niloticus'))
```




```{r}
## map of exports
tmap_mode(mode = 'view') #toggles to interactive viewing
tm_shape(import_export_sf) + #taking a shape..
  tm_fill(col = 'import_count', 
          palette = "viridis",
          alpha = 0.5,
          popup.vars = c("country" = "country", "count of imports" = "import_count")) + #attributing a fill to that shape
  tm_borders(col = 'black')
```


```{r}
## map of exports
tmap_mode(mode = 'view') #toggles to interactive viewing
tm_shape(import_export_sf) + #taking a shape..
  tm_fill(col = 'export_count', 
          palette = "viridis",
          alpha = 0.5,
          popup.vars = c("country" = "country", "count of exports" = "export_count")) + #attributing a fill to that shape
  tm_borders(col = 'black')

```


### Wrangling for Widget 4- Imports and Exports

```{r}
count <- wildlife %>% 
  count(taxon, sort = TRUE) %>% 
  slice_max(n, n = 10)
  
# i dont know how to make this a data frame so the top imports are:
#"Crocodylus niloticus","Python reticulatus","Alligator mississippiensis","Macaca fascicularis" 
## I sliced it to only show the top 10? Not sure if this is what we want? 
 
```

### filter for top 5 imported species
```{r}
top5 <- wildlife %>% 
  filter(taxon %in% c("Crocodylus niloticus", "Python reticulatus","Alligator mississippiensis","Crocodylus porosus","Varanus salvator")) 
```

## calculate imports and exports of top 5 species

```{r}
top5_imports <- top5 %>% 
  group_by(importer) %>% 
  count(taxon, sort = TRUE) 

top5_exports <- top5 %>% 
  group_by(exporter) %>% 
  count(taxon, sort = TRUE)
  
```

Thinking about changing widget 4, wrangling for this change

```{r}
top_purpose <- wildlife %>% 
  group_by(purpose) %>% 
  summarize(count = n()) %>% 
  slice_max(count, n = 6)

commercial <- wildlife %>% 
  filter(purpose == "T") %>% 
  group_by(taxon) %>%  
  summarize(count = n()) %>% 
  slice_max(count, n = )

ggplot(data = commercial, aes(y = taxon, x = count)) +
  geom_col()
```

### Wrangling for Widget 2

Graph of traded species and how much is live or dead

```{r}
# graph of live species for top species

live_species <- wildlife %>% 
  group_by(taxon) %>% 
  filter(term == "live")
  
  

ggplot(data = live_species, aes(count(x = taxon))) +
  geom_histogram()
```

### Data wrangling for widget 3

time data of trade terms

### Merge top 3 files

```{r}
elephants <- read_csv(here('data','elephants.csv'))

oryx <- read_csv(here('data','oryx.csv'))

pythons <- read_csv(here('data','python.csv'))

```

```{r}
top3 <- rbind(elephants, oryx, pythons) %>% 
  distinct() %>% 
  clean_names()
```

```{r}
#top 5 elephant terms
elephant_terms <- top3 %>%  
  filter(taxon == "Loxodonta africana") %>% 
  group_by(year,taxon, term) %>% 
  summarize(count = n()) %>% 
  slice_max(count, n = 5)

ggplot(data = elephant_terms, aes(x = year, y = count)) +
  geom_line(aes(color = term)) +
  theme_minimal() +
  labs(title = "Time Series of Top Traded Elephant Products",
       x = "Year", y = "Count of Terms")

```

```{r}
# top 5 phyton terms
python_terms <- top3 %>% 
  filter(taxon =="Python reticulatus") %>% 
  group_by(year, taxon, term) %>% 
  summarize(count = n()) %>% 
  slice_max(count, n = 5)

ggplot(data = python_terms, aes(x = year, y = count)) +
  geom_line(aes(color = term)) +
theme_minimal() +
  labs(title = "Time Series of Top Traded Reticulated Python Products",
       x = "Year",
       y = "Count of Terms")

  
```

```{r}
# top 5 oryx terms
oryx_terms <- top3 %>% 
  filter(taxon == "Oryx dammah") %>% 
  group_by(year, taxon, term) %>% 
  summarize(count = n()) %>% 
  slice_max(count, n = 5)

ggplot(data = oryx_terms, aes(x = year, y = count)) +
  geom_line(aes(color = term)) +
  theme_minimal() +
  labs(title = "Time Series of Top Traded Oryx Products",
       x = "Year", y = "Count of Terms")



```

```{r}
top3_terms <- rbind(elephant_terms, oryx_terms, python_terms) %>% 
  distinct() %>% 
  clean_names()
```




